The Etiology of Cybersecurity

### 2020-09-03 reports/paper_0/main.tex
Implement Francesco+Livia' feedback

\begin{abstract}
	The objective of this research is to lay the foundations for the
	development of a scientific theory that determines (all and only) the possible
	insecure and secure configurations of any abstract system. We claim
	that cybersecurity weaknesses (i.e. errors) are at the beginning of the
	causality chain that leads to cybersecurity attacks.  We, then,
	formulate a scientific (falsifiable by experiments) hypothesis that we
	use to predict all the weaknesses in the architectural design of an
	abstract system. The mathematical formulation of our hypothesis is
	based on a correlation between the epistemological concepts of
	knowledge, beliefs, and information with the engineering concepts of
	requirements, functional architectures, and channels.  Ultimately, our
	hypothesis allows the definition of the cybersecurity of a system as a
	mathematical formula.  We implemented a prototype cybersecurity risk
	assessment tool that, based on our hypothesis, predicts the weaknesses in a
	UML model of a (cyber-physical) system.
\end{abstract}

@@

Currently, the most complete understanding of insecurity issues, is stored into
a network of databases of weaknesses (e.g.
CWE \autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE \autocite{CVE},
NVD \autocite{NIST2020NVD}), and attacks (e.g. CAPEC \autocite{MITRE2020CAPEC},
ATT\&CK \autocite{MITRE2020ATTACK}).  Those insecurity issues can be related to
the violation of one or more requirements (explicit or implicit) in the
specification, design or implementation of a system. The correlation 
between insecurity flaws and cybersecurity requirements has been used to define
standards such as the IEC 62443-1-3 (the Industrial communication networks -
Network and system security -- Part 3-3: System security requirements and
security levels) which defines requirements as ``confidentiality of information
in transit/at-rest''. More generally, the idea of defining cybersecurity requirements
as properties of a system was initially defined in 1970s with the 
CIA triad (Confidentiality, Integrity, Availability) and refined 
over the decades introducing related concepts such as authenticity or
non-repudiation, or introducing new ones such as ``responsibility'' in the
RITE approach (see \autocite{Samonas2014cia} for an overview of the evolution of the CIA triad).
The link between cybersecurity requirements and vulnerabilities is reported 
in the NVD databases by the CVSS \autocite{Mell2007CVSS} scoring system.
The CVSS evaluates of the severity of a vulnerability
by means of different metrics (such as attack complexity and user interaction)
and quantitatively evaluates the impact on the CIA triad.
While cybersecurity requirements, weaknesses, vulnerabilities, and attacks
have been extensively studied and implemented both in academia and industry
to provide tools for the testing or verification of systems, 
no scientific falsifiable theory correlate
cybersecurity requirements to necessary and sufficient conditions (e.g. mitigations) 
to declare a system secure \autocite{Herley2016unfalsifiability}.
Nonetheless, the extensive body of literature have scientific foundations,
for example, providing formal frameworks for the verification of 
properties %either necessary or sufficient conditions
for cybersecurity.
As a driver for our argumentation, we start by reviewing the key concepts
in the cybersecurity domain.

@@

Our literature review shows that most of the definitions relate insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker). In fact, weaknesses become vulnerability if an
attacker exploits them in an attack.  This, however, just shifts the problem of
defining what cybersecurity is as the problem of defining what a dishonest
agent can or cannot do in a system, where an agent is any virtual or physical
entity of the system or using the system (e.g. a device, a software, or a human
being), and dishonesty is not necessarily related to malicious motivation but
also to incompetence or lack of skills. As in the Dolev-Yao attacker model\footnote{The Dolev-Yao
attacker model can be considered as abstract model of an attacker that can be
used for the analysis of protocol specifications.} \autocite{Dolev1983security}, we may
correlate ``being dishonest'' to ``not following the intended behavior/rules''.
In the case of a generic system, a dishonest agent is, therefore, any agent
that doesn't follow the intended behavior (or functionality or logic) of the
system.  For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an incident. This
may lead to the conclusion that the dishonest behaviors of agents cannot be
defined in general (e.g. due to the heterogeneity of agents and systems) and
that huge repository of dishonest behaviors should be kept as definition, such
as CAPEC \autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
to a cybersecurity protocol should be defined as a number of predefined actions as
in
\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few)\footnote{In formal protocol verifiers, in the DY model, the
attacker has a set of hardcoded inference rules that defines how the attacker
deduces new messages based on the observed message exchange.}.  As depicted in
Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
say that the presence of vulnerabilities is a necessary condition in order to
cause an attack in the system.  Those vulnerabilities are, in turn, caused by
the presence of weaknsesses in the system.  Weaknesses are errors in the design
or implementation of a system.  Therefore, a theory on cybersecurity should
first predict the errors in a system design.

\section{A Cybersecurity Hypothesis in the $\abftheory$-framework}\label{sec:hypothesis}
In order to address the problem raised by Herley, we shall define how to
distinguish between a secure and an insecure system. While most of the
literature have correlated the problem of in-security
to the maliciousness of agents interacting with the system, we show that
cybersecurity doesn't seem to stem from a malicious nature but, rather, insecurity
raises from the lack of well-defined cybersecurity requirements for the development
process of a system. We argue that the high number of cybersecurity vulnerability
reported today are simply the realization of potential system configurations,
which deviate from the nominal behavior because the \emph{intended (or
nominal) behavior} of the system is not precisely defined in the specification
at the very early stages of the engineering process.
As a reference, we consider the following steps of the engineering process
and our objective is to be able to predict the cybersecurity requirements
instead of axiomatically assume them.
\begin{enumerate}
	\item \emph{System Specification}, where the \emph{functional and physical requirements} are defined
	\item \emph{Architecture Design}, where the specification is structured into \emph{functional and physical architectures}
	\item \emph{Cybersecurity Risk Assessment}, where the potential \emph{weaknesses} (errors)
		are identified and, consequently, \emph{cybersecurity requirements} are
		captured.
	%\item \emph{Design Refinement}, where the physical and functional
	%	architectures are extended with \emph{mitigations} for the
	%	weaknesses. Mitigations at this step details the semantics of
	%	the functional blocks (behaviors), the data flow, and communication logic.
	%\item \emph{Verification of Security Requirements}, where the data flow, behaviors, and communication protocols are verified against the cybersecurity requirements.
\end{enumerate}
Given that, in our investigation, we changed the focus from the attacker to the
potential designs of a system, we start by defining a general framework for the definition
of a system. On top of this framework we will identify system weaknesses as potential
design errors.

@@

Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent

@@

proposition known by an agent, such that: (i) knowledge requires belief, (ii)
knowledge require truth, (iii) knowledge must be \emph{properly justified}, and

@@

with an agent's beliefs. Similarly, we consider beliefs to define the actual

@@

		recipient; and we call it \emph{assertion}  

@@

		(output) the behavior of an agent 

@@

		knowledge. Specifically, facts describes: the functional

@@

		subsystem

@@

Similarly to \autocite{Santaca2016abf}, we define agents as a meronomy (an
hierarchy of Part-Whole relations) over the constituent that we previously defined
(assertions, beliefs, and facts), based on a standard definition of mereology,
i.e. based on the definition of parthood relation between \emph{parts}.  Due to
the necessity of considering different relations between parts (as we'll show afterwards)
we extend the mereology to a
mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Table~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{region} both to refer to a mereological part
and to a topological region.  Our aim is to create a meronomy (hierarchy of part-whole
relations)
instead of the taxonomies (categorization based on discrete sets) 
such as the one provided
in \autocite{NIST2020NVD,MITRE2020CVE} 
so that we don't need to rely on a scoring system (such as the 
CVSS) to assign a quantitative evaluation 
of the cybersecurity of each entry. Instead, we want a precise calculation
of the number of insecure configurations to emerge from the mathematical
formulation of our cybersecurity hypothesis.

@@

In Table~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
over the mereotopology using the RCC calculus, as follows, where $\rcc(X,Y)$ on
two generic regions $X,Y$ represents one of the possible RCC relations between
$X$ and $Y$. We note that all the RCC relations are symmetric with the exception
of those that has an explicit (related) inverse.

@@

As in \autocite{Santaca2016abf}, it follows that, defining
a system with a fixed number of regions, there exist
an upper-bound to the number of possible configuration of a system, defined by
the possible relations between the different regions.
For completeness, we report in the next paragraph 
the calculation done by the authors in \autocite{Santaca2016abf}.

@@

\begin{itemize}%[nosep]
\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15 
\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54
\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193
\end{itemize}

Hence, even if considering a different number of regions than the three
$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
the number of theoretical agents, the application of RCC downscales that number
of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
3.6 times more (different) types of agents than RCC3, but using RCC8 would
allow us to consider 3.5 times more different agents.
In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior. We note 
that, the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system.
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{quantitative.pdf}
	\caption{Cybersecurity risk for a single agent}
	\label{fig:quantitative}
\end{figure}

@@

\paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and
\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} 
The relation of assertions
and beliefs with facts determines the quality with respect to the nominal
(specified) system.  Given that facts defines what needs to be true in the system,
the relation of assertions an facts determines the degree of quality between
the real information circulating in a system (or within an agent) and
the one specified.  Since the transfer of information
through assertions generates beliefs, a dishonest agent may circulate false
information, generating false beliefs.  We note that it is often implied that
the intention behind circulating false information discerns a dishonest and an
incompetent agent. However, we consider honesty related to sharing
truths.
%\footnote{``[\ldots] truth exists only in the good man, but the true in
%the bad man as well; for it is possible for the bad man to utter something
%true.'' Sextus Empiricus, Outlines of Pyrrhonism,
%II-83 \autocite{Empiricus1990Pyrrhonism}}
The relation between beliefs and
facts determines the competence (on the subjects defined by the facts) of an
agent (i.e. the more competent an agent is, the more likely a belief of that
agent is true).

@@

\section{Prediction of Cybersecurity Weaknesses}\label{sec:theory}
Several standards mandate a secure-by-design approach in which cybersecurity
shall be considered at the very early stages of the design process.  For
example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and
		``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
Standards do not describe in detail how to perform a cybersecurity risk 
assessment and only vaguely define the overall objective, which
can be summarized as to provide an understanding of the potential cybersecurity risks.
Roughly speaking, a cybersecurity risk assessment (e.g. CORAS \autocite{Lund2010model}) 
methodology starts from the (i) identification of assets, (ii) determines
the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
threat diagrams in the CORAS
approach, (iii) calculates the risk as a function
of impact and likelihood of threat scenarios, and (iv) finally provides
treatments (mandating a partial re-design) and residual risks.
All the methodologies and tools we reviewed (e.g. Threatmodeler \autocite{Threatmodeler},
CORAS \autocite{Lund2010model}, SECRAM \autocite{De2015role})
rely on the expertise of the person who performs the risk assessment for
the identification of threats and for the quantitative estimation of risks.
As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
we don't have a metric for security'' while standards require a cybersecurity risk assessment
which, in turn, mandates a metric to evaluate the cybersecurity risk. 
In contrast, in this section, we define how to specify a CPS in the $\abftheory$-framework
and we identify a cybersecurity metric.

@@

\paragraph{Port Weaknesses} We now apply the $\abftheory$-framework to
list all possible cybersecurity weaknesses of a port.
\begin{theorem}{Port Weaknesses.}
From our definition of ports, it follows that there exist only the following \emph{six}
types of weaknesses, generating six types of insecure port in RCC5:
%(the notation is reported for input-ports, i.e. on the left-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{replace port}, where assertions reach
		the port but is replaced with different and un-related information
		before passing the boundary 
	\item \emph{drop port}, where assertions reach the
		port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{insertion port}, where some new information is transferred along with 
		the information incoming from a channel, and then sent to
		the intended recipient (agent)
	\item \emph{injection port}, where part of the incoming information
		is substituted with new information and transferred to the intended recipient
	\item \emph{selective port}, where some information passes the port and
		part is either:
	\begin{enumerate}[start=1, label={W5.\arabic*)}]
		\item dropped, 
		\item replaced.
	\end{enumerate}
\end{enumerate}
\end{theorem}

\begin{proof}
An input port is, in the $\abftheory$-framework, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different types
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Let us
	define a function weight $|X|$ such that, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every open region can be written as union of elements of that base. 
	We distinguish between regions that contain information and regions that
	don't by writing the latter as $\emptyset$.
	\begin{enumerate}
		\item if $\eq{\assertionRegion}{\beliefRegion}$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication)
		\item if $\dr{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former and \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			which we call \emph{replace} (i.e. drop and insert)
		\item if $\pp{\assertionRegion}{\beliefRegion}$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{insertion}
		\item if $\ppi{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{drop} (or selective drop to stress the difference with the full drop)
		\item if $\po{\assertionRegion}{\beliefRegion}$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation} which we call \emph{injection}.
	\end{enumerate}
\end{proof}

@@

\begin{enumerate}[start=6, label={W\arabic*)}]
	\item \emph{secure output port} and \emph{input drop port}
	\item \emph{secure output port} and \emph{input insertion port}
	\item \emph{output drop port} and \emph{input drop port}
	\item \emph{output drop port} and \emph{input insertion port}
	\item \emph{output drop port} and \emph{input secure port}
	\item \emph{output injection port} and \emph{input secure port}
\end{enumerate}

>>>>

\begin{abstract}
	The objective of this research is to lay the foundations for the
	development of a scientific theory that determines (all and only) the possible
	insecure and secure configurations of any abstract system. We claim
	that cybersecurity weaknesses (i.e. errors) are at the beginning of the
	causality chain that leads to cybersecurity attacks.  We, then,
	formulate a scientific (falsifiable by experiments) hypothesis that we
	use to predict all the weaknesses in the architectural design of an
	abstract system. The mathematical formulation of our hypothesis is
	based on a correlation between the epistemological concepts of
	knowledge, beliefs, and information with the engineering concepts of
	requirements, functional architectures, and channels.  Ultimately, our
	hypothesis allows for the definition of the cybersecurity of a system as a
	mathematical formula.  We implemented a prototype cybersecurity risk
	assessment tool that, based on our hypothesis, predicts the weaknesses in a
	UML model of a (cyber-physical) system.
\end{abstract}

@@

Currently, the most complete understanding of insecurity issues is stored into
a network of databases of weaknesses (e.g.
CWE \autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE \autocite{CVE},
NVD \autocite{NIST2020NVD}), and attacks (e.g. CAPEC \autocite{MITRE2020CAPEC},
ATT\&CK \autocite{MITRE2020ATTACK}).  Those insecurity issues can be related to
the violation of one or more requirements (explicit or implicit) in the
specification, design or implementation of a system. The correlation 
between insecurity flaws and cybersecurity requirements has been used to define
standards such as the IEC 62443-1-3 (the Industrial communication networks -
Network and system security -- Part 3-3: System security requirements and
security levels) which defines requirements as ``confidentiality of information
in transit/at-rest''. More generally, the idea of defining cybersecurity requirements
as properties of a system was initially defined in 1970s with the 
CIA triad (Confidentiality, Integrity, Availability) and refined 
over the decades introducing related concepts such as authenticity or
non-repudiation, or introducing new ones such as ``responsibility'' in the
RITE approach (see \autocite{Samonas2014cia} for an overview of the evolution of the CIA triad).
The link between cybersecurity requirements and vulnerabilities is reported 
in the NVD databases by the CVSS \autocite{Mell2007CVSS} scoring system.
The CVSS evaluates of the severity of a vulnerability
by means of different metrics (such as attack complexity and user interaction)
and quantitatively evaluates the impact on the CIA triad.
While cybersecurity requirements, weaknesses, vulnerabilities, and attacks
have been extensively studied and implemented both in academia and industry
to provide tools for the testing or verification of systems, 
no scientific falsifiable theory correlates
cybersecurity requirements to necessary and sufficient conditions (e.g. mitigations) 
to declare a system secure \autocite{Herley2016unfalsifiability}.
Nonetheless, the extensive body of literature has scientific foundations,
for example, providing formal frameworks for the verification of 
properties %either necessary or sufficient conditions
for cybersecurity.
As a driver for our argumentation, we start by reviewing the key concepts
in the cybersecurity domain.

@@

Our literature review shows that most of the definitions relate insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker). In fact, weaknesses become vulnerability if an
attacker exploits them in an attack.  This, however, just shifts the problem of
defining what cybersecurity is as the problem of defining what a dishonest
agent can or cannot do in a system, where an agent is any virtual or physical
entity of the system or using the system (e.g. a device, a software, or a human
being), and dishonesty is not necessarily related to malicious motivation but
also to incompetence or lack of skills. As in the Dolev-Yao attacker model\footnote{The Dolev-Yao
attacker model can be considered as abstract model of an attacker that can be
used for the analysis of protocol specifications.} \autocite{Dolev1983security}, we may
correlate ``being dishonest'' to ``not following the intended behavior/rules''.
In the case of a generic system, a dishonest agent is, therefore, any agent
that doesn't follow the intended behavior (or functionality or logic) of the
system.  For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an incident. This
may lead to the conclusion that the dishonest behaviors of agents cannot be
defined in general (e.g. due to the heterogeneity of agents and systems) and
that a huge repository of dishonest behaviors should be kept as definition, such
as CAPEC \autocite{MITRE2020CAPEC}; or that dishonesty of an agent with respect
to a cybersecurity protocol should be defined as a number of predefined actions as
in
\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few)\footnote{In formal protocol verifiers, in the DY model, the
attacker has a set of hardcoded inference rules that defines how the attacker
deduces new messages based on the observed message exchange.}.  As depicted in
Figure~\ref{fig:causality}, in order to define a theory on cybersecurity we may
say that the presence of vulnerabilities is a necessary condition in order to
cause an attack in the system.  Those vulnerabilities are, in turn, caused by
the presence of weaknsesses in the system.  Weaknesses are errors in the design
or implementation of a system.  Therefore, a theory on cybersecurity should
first predict the errors in a system design.

\section{A Cybersecurity Hypothesis in the $\abftheory$-framework}\label{sec:hypothesis}
In order to address the problem raised by Herley, we shall define how to
distinguish between a secure and an insecure system. While most of the
literature have correlated the problem of in-security
to the maliciousness of agents interacting with the system, we show that
cybersecurity doesn't seem to stem from a malicious nature but, rather, insecurity
raises from the lack of well-defined cybersecurity requirements for the development
process of a system. We argue that the high number of cybersecurity vulnerability
reported today are simply the realization of potential system configurations,
which deviate from the nominal behavior because the \emph{intended (or
nominal) behavior} of the system is not precisely defined in the specification
at the very early stages of the engineering process.
As a reference, we consider the following steps of the engineering process
and our objective is to be able to predict the cybersecurity requirements
instead of axiomatically assume them.
\begin{enumerate}
	\item \emph{System Specification}, where the \emph{functional and physical requirements} are defined.
	\item \emph{Architecture Design}, where the specification is structured into \emph{functional and physical architectures}.
	\item \emph{Cybersecurity Risk Assessment}, where the potential \emph{weaknesses} (errors)
		are identified and, consequently, \emph{cybersecurity requirements} are
		captured.
	%\item \emph{Design Refinement}, where the physical and functional
	%	architectures are extended with \emph{mitigations} for the
	%	weaknesses. Mitigations at this step details the semantics of
	%	the functional blocks (behaviors), the data flow, and communication logic.
	%\item \emph{Verification of Security Requirements}, where the data flow, behaviors, and communication protocols are verified against the cybersecurity requirements.
\end{enumerate}
Given that, in our investigation, we changed the focus from the attacker to the
potential designs of a system, we start by defining a general framework for the definition
of a system. On top of this framework we will identify system weaknesses as potential
design errors.

@@

Multi-Agent-System, Epistemic Logic) on what the constituent of an agent are,

@@

propositions known by an agent, such that: (i) knowledge requires belief, (ii)
knowledge requires truth, (iii) knowledge must be \emph{properly justified}, and

@@

to an agent's beliefs. Similarly, we consider beliefs to define the actual

@@

		recipient; and we call it \emph{assertion}.

@@

		(output) the behavior of an agent.

@@

		knowledge. Specifically, facts describe: the functional

@@

		subsystem.

@@

Similarly to \autocite{Santaca2016abf}, we define agents as a meronomy (an
hierarchy of Part-Whole relations) over the constituents that we previously defined
(assertions, beliefs, and facts), based on a standard definition of mereology,
i.e. based on the definition of parthood relation between \emph{parts}.  Due to
the necessity of considering different relations between parts (as we'll show afterwards)
we extend the mereology to a
mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Table~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{region} both to refer to a mereological part
and to a topological region.  Our aim is to create a meronomy (hierarchy of part-whole
relations)
instead of the taxonomies (categorization based on discrete sets) 
such as the one provided
in \autocite{NIST2020NVD,MITRE2020CVE} 
so that we don't need to rely on a scoring system (such as the 
CVSS) to assign a quantitative evaluation 
of the cybersecurity of each entry. Instead, we want a precise calculation
of the number of insecure configurations to emerge from the mathematical
formulation of our cybersecurity hypothesis.

@@

In Table~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
over the mereotopology using the RCC calculus, as follows, where $\rcc(X,Y)$ on
two generic regions $X,Y$ represents one of the possible RCC relations between
$X$ and $Y$. We note that all the RCC relations are symmetric with the exception
of those that have an explicit (related) inverse.

@@

As in \autocite{Santaca2016abf}, it follows that, defining
a system with a fixed number of regions, there exists
an upper-bound to the number of possible configuration of a system, defined by
the possible relations between the different regions.
For completeness, we report in the next paragraph 
the calculation done by the authors in \autocite{Santaca2016abf}.

@@

\begin{itemize}%[nosep]
\item \emph{RCC3} --- theoretical: $3^3=27$,  correct: 15.
\item \emph{RCC5} --- theoretical: $5^3=125$, correct: 54.
\item \emph{RCC8} --- theoretical: $8^3=512$, correct: 193.
\end{itemize}

Hence, even if considering a different number of regions than the three
$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
the number of theoretical agents, the application of RCC downscales that number
of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
3.6 times more (different) types of agents than RCC3, but using RCC8 would
allow us to consider 3.5 times more different agents.
In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior. We note 
that the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system.
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{quantitative.pdf}
	\caption{Cybersecurity risk for a single agent}
	\label{fig:quantitative}
\end{figure}

@@

\paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and
\Rcc{$\beliefRegion$}{$\factRegion$} -- Honesty and Competence} 
The relation of assertions
and beliefs with facts determines the quality with respect to the nominal
(specified) system.  Given that facts define what needs to be true in the system,
the relation of assertions and facts determines the degree of quality between
the real information circulating in a system (or within an agent) and
the one specified.  Since the transfer of information
through assertions generates beliefs, a dishonest agent may circulate false
information, generating false beliefs.  We note that it is often implied that
the intention behind circulating false information discerns a dishonest and an
incompetent agent. However, we consider honesty related to sharing
truths.
%\footnote{``[\ldots] truth exists only in the good man, but the true in
%the bad man as well; for it is possible for the bad man to utter something
%true.'' Sextus Empiricus, Outlines of Pyrrhonism,
%II-83 \autocite{Empiricus1990Pyrrhonism}}
The relation between beliefs and
facts determines the competence (on the subjects defined by the facts) of an
agent (i.e. the more competent an agent is, the more likely a belief of that
agent is true).

@@

\section{Prediction of Cybersecurity Weaknesses}\label{sec:theory}
Several standards mandate a secure-by-design approach in which cybersecurity
shall be considered at the very early stages of the design process.  For
example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and
		``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''.
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''.
\end{enumerate}
Standards do not describe in detail how to perform a cybersecurity risk 
assessment and only vaguely define the overall objective, which
can be summarized as to provide an understanding of the potential cybersecurity risks.
Roughly speaking, a cybersecurity risk assessment (e.g. CORAS \autocite{Lund2010model}) 
methodology starts from the (i) identification of assets, (ii) determines
the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
threat diagrams in the CORAS
approach, (iii) calculates the risk as a function
of impact and likelihood of threat scenarios, and (iv) finally provides
treatments (mandating a partial re-design) and residual risks.
All the methodologies and tools we reviewed (e.g. Threatmodeler \autocite{Threatmodeler},
CORAS \autocite{Lund2010model}, SECRAM \autocite{De2015role})
rely on the expertise of the person who performs the risk assessment for
the identification of threats and for the quantitative estimation of risks.
As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
we don't have a metric for security'' while standards require a cybersecurity risk assessment
which, in turn, mandates a metric to evaluate the cybersecurity risk. 
In contrast, in this section, we define how to specify a CPS in the $\abftheory$-framework
and we identify a cybersecurity metric.

@@

\paragraph{Port Weaknesses} We now apply the $\abftheory$-framework to
list all possible cybersecurity weaknesses of a port.
\begin{theorem}{Port Weaknesses.}
From our definition of ports, it follows that there exist only the following \emph{six}
types of weaknesses, generating six types of insecure port in RCC5:
%(the notation is reported for input-ports, i.e. on the left-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{Replace port}, where assertions reach
		the port but is replaced with different and un-related information
		before passing the boundary.
	\item \emph{Drop port}, where assertions reach the
		port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent).
	\item \emph{Insertion port}, where some new information is transferred along with 
		the information incoming from a channel, and then sent to
		the intended recipient (agent).
	\item \emph{Injection port}, where part of the incoming information
		is substituted with new information and transferred to the intended recipient.
	\item \emph{Selective port}, where some information passes the port and
		part is either:
	\begin{enumerate}[start=1, label={W5.\arabic*)}]
		\item Dropped, or 
		\item Replaced.
	\end{enumerate}
\end{enumerate}
\end{theorem}

\begin{proof}
An input port is, in the $\abftheory$-framework, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different types
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Let us
	define a function weight $|X|$ such that, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every open region can be written as union of elements of that base. 
	We distinguish between regions that contain information and regions that
	don't by writing the latter as $\emptyset$.
	\begin{enumerate}
		\item If $\eq{\assertionRegion}{\beliefRegion}$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication).
		\item If $\dr{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former and \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			which we call \emph{replace} (i.e. drop and insert).
		\item If $\pp{\assertionRegion}{\beliefRegion}$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{insertion}.
		\item If $\ppi{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{drop} (or selective drop to stress the difference with the full drop).
		\item If $\po{\assertionRegion}{\beliefRegion}$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation} which we call \emph{injection}.
	\end{enumerate}
\end{proof}

@@

\begin{enumerate}[start=6, label={W\arabic*)}]
	\item \emph{Secure output port} and \emph{input drop port}.
	\item \emph{Secure output port} and \emph{input insertion port}.
	\item \emph{Output drop port} and \emph{input drop port}.
	\item \emph{Output drop port} and \emph{input insertion port}.
	\item \emph{Output drop port} and \emph{input secure port}.
	\item \emph{Output injection port} and \emph{input secure port}.
\end{enumerate}
